---
title: "Introduction to simesHotelling"
author: "Tzviel Frostig"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(simesHotelling)
library(MASS)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Introduction 

The package contains the different variants of the SH test, allowing to test high-dimensional mean vectors for when $p > n$. 
The tests available in the packge are: Parametric SH, non-equal covariance SH, and non-parametric SH test (permutation test). 


#### Parametric tests 


Creating the sample data $n_X = n_y = 50, p = 100, \Sigma_X = \Sigma_Y = I, \mu_x = \vec{0}, \mu_y = \vec{0}$. 

##### One sample SH 

```{r}
library(MASS)
X <- mvrnorm(50, rep(0, 100), diag(100)) 
```

One sample SH test where $m = 30, B = 100$ and $\delta = \vec{0}$

```{r}
SHTest(X = X,
       samp.size = 30,
       iterations = 100,
       delta = rep(0, 100))
```

It returns the p-value of the test, the F-statistics of the each group of sampled dimensions and the p-value of teh resulting F test for each group of sampled dimensions. 

##### Randomness of test 
Notice that the test is random so each iterations will yield a slightly different p-value unless you use set.seed. 

```{r}
SHTest(X = X,
       samp.size = 30,
       iterations = 100,
       delta = rep(0, 100))$`P-value`

set.seed(999)
SHTest(X = X,
       samp.size = 30,
       iterations = 100,
       delta = rep(0, 100))$`P-value`
set.seed(999)
SHTest(X = X,
       samp.size = 30,
       iterations = 100,
       delta = rep(0, 100))$`P-value`

```


##### Two sample SH 
It can also be used for testing the difference in the mean vector of X and Y. $Y$ is generated from the same distribution as $X$.

```{r}
Y <- mvrnorm(50, rep(0, 100), diag(100)) 

SHTest(X = X,
       Y = Y, 
       samp.size = 30,
       iterations = 100,
       delta = rep(0, 100))
```

###### Two sample SH for non-equal covariance matrices  

As well as testing the difference in the mean vector why X and Y have different covariance matrices. 

Generating data $n_X = 50, n_Y = 79, \mu_X = \mu_Y = 0, \Sigma_X = I, \Sigma_Y = ar.mat$. 
```{r}
ar.mat <- 0.8^abs(outer(1:100, 1:100, "-"))
### Generating data 
X <- mvrnorm(50, rep(0, 100), diag(100)) 
Y <- mvrnorm(79, rep(0, 100), ar.mat)


SHTest(X, 
       Y, 
       samp.size = 40, 
       iterations = 100, 
       equal.cov = FALSE,
       delta = rep(0, 100))
```



#### Non - Parametric tests 

It is all possible to conduct a non-parametric test. We will describe here the workflow. 

Generating data $n_X = 50, n_Y = 100, p = 200, \Sigma = I$. 
```{r}
X <- mvrnorm(50, rep(0, 200), diag(200)) 
Y <- mvrnorm(100, rep(0, 200), diag(200))
```

Preparing the data, how many permutation should be conducted, the number of dimensions sampled and the number of sampling to conduct.  

```{r}
L <- PrepareNonPara(X, Y,
               permutation.num = 100, 
               iter.num = 100, 
               samp.size = 50) 
```

The functions returns a large list including the original X, Y, a list of permuted X and Y, this is done to allow comparisions to different permutations tests, and the samp.mat which is the matrix indicating how the samples should be done. 
Now the non-parametric test can be conducted, two options are available, 'SH' based on Simes and Thullin (see A high-dimensional two-sample test for the mean using random subspaces
- Mans Thulin)

```{r}
SHnonPara(L, method = 'SH')
SHnonPara(L, method = 'Thulin')
```

